===---
title: "Lab-10_28-4-22"
author: "21122018 Drashty Ranpara"
date: "30/04/2022"
output: html_document
---
#Importing the Dataset.
```{r}
library(readxl)
a=read_excel("C:\\Users\\Lenovo\\Desktop\\MscDs\\SEM-2\\Programming in R\\energydata_complete.xlsx")
```

#Viewing the Dataset.
```{r}
View(a)
```

#Structure of the Dataset.
```{r}
str(a)
```
#Column Names of the Dataset.
```{r}
names(a)
```
#Dimensions of the Dataset.
```{r}
dim(a)
```
#Summary of the Dataset.
```{r}
summary(a)
```
#First 5 Rows from the Dataset.
```{r}
head(a)
```

#Last 5 Rows from the Dataset.
```{r}
tail(a)
```

#Class of the Dataset.
```{r}
class(a)
```
#Minimum, Maximum and Range of only Appliances Variable.
```{r}
min(a[,2])
max(a[,2])
range(a[,2])
```
#Quantiles of all the features except date feature.
```{r}
apply(a[,2:29],2,quantile)
```
#Standard Deviation of all the variables exclusing Date features.
```{r}
apply(a[,2:29], 2, sd)
```
#Correlation between all the numerical features of the dataset. 
```{r}
cor(a[2:28])
```
#EDA-Exploratory Data Analysis.

#Heatmap or Correlation Plot.
```{r}
library(corrplot)
corrplot(cor(a[2:29]), method = "circle")
```

#Sample BarPlot of the first 100 rows of the Appliances Feature.
```{r}
library(RColorBrewer)
barplot(head(a$Appliances,100),col  = brewer.pal(3,"Set1"),xlab="Appliances",main="Barplot of Appliances")
```

#Histogram of the first 500 rows of the Lights Column.
```{r}
hist(head(a$lights,500), col = "red",xlab="Lights",ylab="Frequency Plotting of Lights",main="Histogram of Lights")
```
#QPlot for the Date and Appliances Variables.
```{r}
require(gridExtra)
library(ggplot2)
plot1 <-qplot(a$date,a$Appliances,xlab='Date',ylab='Appliances Wh',geom="line",col="brown")+theme_grey(base_size = 18) 
plot1
```
#Boxplot of the Appliances Column.
```{r}
boxplot(a$Appliances,
        boxfill = "lightblue",horizontal=TRUE,ylim=c(0,1200),xlab="Appliances Wh",frame=F,
        cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
```
#Boxplot of the Lights Column.
```{r}
boxplot(a$lights,
        boxfill = "darkblue",horizontal=TRUE,ylim=c(0,1200),xlab="Appliances Wh",frame=F,
        cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
```
#GGplot for Lights and Appliances Variables.
```{r}
ggplot(a,aes(x = lights, y = Appliances)) + geom_point() +geom_smooth(method = "lm")
```

#GGPlot of Windspeed and Visibility Variables.
```{r}
ggplot(a,aes(x = Windspeed, y = Visibility)) + geom_point() +geom_smooth(method = "lm")
```
#ScatterPlot of Temperature and Humidity for Kitchen Area.
```{r}
library(scatterplot3d)
scatterplot3d(a$T1,a$RH_1,
              pch=16,type="h",
              angle=45,
              xlab="Temperature",
              ylab="Humidity",
              main="Scatter Plot of Temperature and Humidity of Kitchen Area",
              color=c("purple"))
```
#ScatterPlot of Temperature and Humidity for Office Room.
```{r}
library(scatterplot3d)
scatterplot3d(a$T4,a$RH_4,
              pch=16,type="h",
              angle=45,
              xlab="Temperature in Celsius",
              ylab="Humidity",
              main="Scatter Plot of Temperature and Humidity of Office Room",
              color=c("brown"))
```
#KDE Plot of Windspeed.
```{r}
plot(a$Windspeed, main="Kernel Density of WindSpeed ")
polygon(a$Windspeed, col="red", border="blue")
```
#KDE Plot of Pressure from Weather Station in mm hg.
```{r}
den <- density(a$Press_mm_hg)
plot(den, frame = FALSE, col = "orange",main = "Density plot of Pressure from Weather Station")
```
#Histogram of Temperature in Teenage Room.
```{r}
hist(a$T8, 
     col="chocolate3",
     border="black",
     prob = TRUE,
     xlab = "Tempertaure in Celsius",
     main = "Temperature in Teenage Room")
  
lines(density(a$T8),
      lwd = 2,
      col = "black")
```
#Sample Pie Chart for first 20 rows for date and Appliances columns.
```{r}
pie(head(a$Appliances,20), head(a$date,20), main = "Appliances pie chart", col = rainbow(length(a$Appliances)))
```
#Multiple Linear Regression Model of rv1 with all the columns.
```{r}
linearm=lm(rv1~., data = a[2:28])
linearm
```
#Summary of Multiple Linear Regression Model.
```{r}
summary(linearm)
```
#Predicting the rv1 with linear regression model and storing it into predscore column which will be added in the dataset..
```{r}
a$predscore=predict(linearm)
View(a)
```

```{r}
a
```

#R Square Score of the Multiple Linear Regression Model.
```{r}
summary(linearm)$r.squared
```
#The r2 core is very low i.e. 0.12% and hence the model is not a good fit for the data.

