---
title: "R Notebook"
output: html_notebook
---
## Take airquality dataset
```{r}
df=airquality
df
```

## Replacing missing values with mean
```{r}
#library(zoo); data3 <- na.aggregate(df); colSums(data3) 
data=df
for(i in 1:ncol(df)) {                                   # Replace NA in all columns
  data[ , i][is.na(data[ , i])] <- mean(data[ , i], na.rm = TRUE)
}
colSums(data)
```
## Display randomForest model
```{r}
set.seed(1)
library(randomForest)
model = randomForest(Ozone~., data)
model
```
## Plot the model
```{r}
plot(model)
```
# Find minimum MSE
```{r}
which.min(model$mse)
```
## Gives minimum RMSE value
```{r}
sqrt(model$mse[which.min(model$mse)])
```

## Produces variable importance plot
```{r}
varImpPlot(model)
```

## Random train test split
```{r}
size=floor(0.80*nrow(data))
# generate random sample of dataset
set.seed(99)
indexes = sample(seq_len(nrow(data)),size)
train <- data[indexes, ]
test <- data[-indexes, ]
summary(train)
summary(test)
```
## Run training on random forest classifier
```{r}
# Create a Random Forest model with default parameters
model1 <- randomForest(Ozone ~ ., data = train, importance = TRUE)
model1
```
## Fine tuning parameters of Random Forest model
```{r}
model2 <- randomForest(Ozone ~ ., data = train, ntree = 500, mtry = 6, importance = TRUE)
model2
```
## Predicting on train set
```{r}
predTrain <- predict(model2, train, type = "class")
```
## Checking classification accuracy
```{r}
table(predTrain, train$Ozone) 
```
## Predicting on Validation set
```{r}
predValid <- predict(model2, test, type = "class")
```
## Checking classification accuracy
```{r}
mean(predValid == test$Ozone)                    
table(predValid,test$Ozone)
```
## To check important variables
```{r}
importance(model2)        
varImpPlot(model2)
```
## Now, we will use ‘for’ loop and check for different values of mtry.
```{r}
# Using For loop to identify the right mtry for model
a=c()
i=5
for (i in 3:8) {
  model3 <- randomForest(Condition ~ ., data = train, ntree = 500, mtry = i, importance = TRUE)
  predValid <- predict(model3, test, type = "class")
  a[i-2] = mean(predValid == test$Ozone)
}
a
```

